## Chapter 4 - Discovering Statistics Using R

# I'm using the data from the personalitytest to try all the different functions introduced throughout the chapter

```{reading the data frame}
df <- read_csv("NEW_CogSciPersonalityTest2019 (1).csv")
```

# I get tidyverse into my library

```{r}
library(tidyverse)
```

#I tell ggplot which dataframe to use:

```{r}
ggplot(df)
```

# I group something haha

```{r}
summary_gender_mean_showsize <- df %>%
  group_by(gender) %>%
summarize(mean(shoesize))
```

# I tell ggplot to make a graph with gender on the x-axis and the mean showsize on the y-axis)

```{r}
ggplot(df, aes(x=gender, y=shoesize)) + geom_bar(stat='summary', fun.y = mean)
```

### Gender and Balloon Balance

# I want to make a histogram that shows the mean of time balancing a balloon according to sex 

```{r}
ggplot(df, aes(x=gender, y=balloon_balance)) + geom_bar(stat='summary', fun.y = mean)
```

# I just did this to check if the mean values were correct:

```{r}
gender_balloon_balance <- df %>%
  group_by(gender) %>%
  summarise(mean(balloon_balance))
```

# Fortunately it was correct

# Now I want the bars to have different colours:

```{r}
ggplot(df, aes(x=gender, y=balloon_balance, fill = gender))+
  geom_bar(stat='summary', fun.y = mean, width = 0.5)+
  labs(x = "Gender", y = "Balloon time")+
  theme_minimal()
```
